---
title: Women significantly outnumber men in Hong Kong
author: Chris Peralta
date: '2017-10-20'
slug: hk-gender
categories:
  - R
tags:
  - ggplot2
  - data visualization
  - design
description: ''
#output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Women significantly outnumber men in Hong Kong. In this article, I will show you how badly men are outnumbered and how significant the gap is. 

## How badly do women outnumber men?

Women make up `r percent(sum(filter(pop, sex == 'female')$population, na.rm = TRUE) / sum(pop$population, na.rm = TRUE))` of the population, while men make up `r percent(1 - sum(filter(pop, sex == 'female')$population, na.rm = TRUE) / sum(pop$population, na.rm = TRUE))` of the population. 

This might not seem so bad, but I'm going to break it down to give you a better picture. First lets look at a population pyramid so we can get an idea of what Hong Kong's population looks like. 

```{r cars}
#population pyramid
```

From this we see that people from 30-59 make up most of the population of Hong Kong with men being older on average than women. 

Now, I'll put the percent of women in each age group next to the percent of men in each age group. 

```{r pressure, echo=FALSE}
#population by age and gender
```

First, lets look at some of the individual points

```{r }
#highlight ends
```

These points are quite interesting, but they can be easily explained.
For the left side:
More boys than girls are born globally, you can read about this phenomena here. INSERT HYPERLINK
For the right side:
Women have significantly longer life spans. 

So from here we will only look at the data for 20-59 year olds as shown below.

```{r }
#highlight target
#include bar chart in upper right corner showing %women and men
```

The 35-39 age group alone, is made up of 60.1% women! This is also one of the largest age groups in Hong Kong!

In the 20-59 age group, there are 56% women to 44% men. This group makes up `r sum(filter(pop, age_range == '20 - 24' | age_range == '25 - 29'| age_range == '30 - 34' | age_range == '35 - 39'| age_range == '40 - 44' | age_range == '45 - 49'| age_range == '50 - 54' | age_range == '55 - 59')$population, na.rm = TRUE) / sum(pop$population, na.rm = TRUE)` of the total population of Hong Kong. 

In this group there are `r sum(filter(pop, (sex == 'female') & (age_range == '20 - 24' | age_range == '25 - 29'| age_range == '30 - 34' | age_range == '35 - 39'| age_range == '40 - 44' | age_range == '45 - 49'| age_range == '50 - 54' | age_range == '55 - 59'))$population, na.rm = TRUE) - sum(filter(pop, (sex == 'male') & (age_range == '20 - 24' | age_range == '25 - 29'| age_range == '30 - 34' | age_range == '35 - 39'| age_range == '40 - 44' | age_range == '45 - 49'| age_range == '50 - 54' | age_range == '55 - 59'))$population, na.rm = TRUE)` more women than men.

You can look at the graph below to see which nationalities contribute the most women to the number above.  

```{r }
#diverging lollipop
```

This graph shows us that the Filipinos and Indonesians contribute almost as many women to the gender gap as the locals of Hong Kong, despite making up a much smaller portion of the overall population. This is likely due to the large numbers of domestic helpers employed by families in Hong Kong, which are almost all women from The Philippines and Indonesia. 











