---
title: Women significantly outnumber men in Hong Kong
author: Chris Peralta
date: '2017-10-20'
slug: hk-gender
categories:
  - R
tags:
  - ggplot2
  - data visualization
  - design
description: ''
#output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Women significantly outnumber men in Hong Kong. In this article, I will show you how badly men are outnumbered and how significant the gap is. 

## How badly do women outnumber men?

Women make up `r percent(sum(filter(pop, sex == 'female')$population, na.rm = TRUE) / sum(pop$population, na.rm = TRUE))` of the population, while men make up `r percent(1 - sum(filter(pop, sex == 'female')$population, na.rm = TRUE) / sum(pop$population, na.rm = TRUE))` of the population. 

This might not seem so bad, but I'm going to break it down to give you a better picture. First lets look at a population pyramid so we can get an idea of what Hong Kong's population looks like. 

```{r cars}
#population pyramid
```

From this we see that people from 30-59 make up most of the population of Hong Kong with men being older on average than women. 

Now, I'll put the percent of women in each age group next to the percent of men in each age group. 

```{r pressure, echo=FALSE}
#population by age and gender
```

First, lets look at some of the individual points

```{r }
#highlight ends
```

These points are quite interesting, but they can be easily explained.
For the left side:
More boys than girls are born globally, you can read about this phenomena here. INSERT HYPERLINK
For the right side:
Women have significantly longer life spans. 

So from here we will only look at the data for 20-59 year olds as shown below.

```{r }
#highlight target
#include bar chart in upper right corner showing %women and men
```

In the 20-59 age group, we have 56% women to 44% men with this group making up `r sum(filter(pop, age_range == '20 - 24' | age_range == '25 - 29'| age_range == '30 - 34' | age_range == '35 - 39'| age_range == '40 - 44' | age_range == '45 - 49'| age_range == '50 - 54' | age_range == '55 - 59')$population, na.rm = TRUE) / sum(pop$population, na.rm = TRUE)` of the population. 

Now lets break this down by nationality to see who contributes the most to this number. 

```{r }
#diverging lollipop
```











